---
title: "Week 9"
author: "DJM"
date: "10 March 2020"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: [default, rladies, rladies-fonts, extra.css]
    nature:
      beforeInit: ["macros.js", "https://platform.twitter.com/widgets.js"]
      titleSlideClass: [center,middle, inverse, bigger]
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
---


# News discussion

.center[.middle[.bigger[
[IU costs too much](https://www.idsnews.com/article/2020/02/opinion-purdue-saves-students-money-while-iu-doesnt?fbclid=IwAR3rTRZ88t1r5C5jfztS4gjWz98T1dFEIHIGb1Xxq2CPRWxuwV54dDnzmbY)
]]]

---

class: inverse, center, middle

# Signal and the Noise (wrapup)


---

# Climate update


```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(cowplot)
global.climate = read_csv("https://www.ncdc.noaa.gov/cag/global/time-series/globe/land_ocean/1/1/1880-2020/data.csv?trend=true&trend_base=10&firsttrendyear=1880&lasttrendyear=2020", skip=4)
base = global.climate$Value[global.climate$Year==1990]
uncertainty = tibble(
  Value = 1990:max(global.climate$Year),
  ymin = 0:(length(Value)-1)*.02, ymax = 0:(length(Value)-1)*.05)
global.climate = global.climate %>% mutate(Value = Value - base) 
```

```{r, fig.align='center', fig.width=14,fig.height=7, echo=FALSE}
ggplot(global.climate, aes(Year,Value)) + 
  ylab('deviation from 1990 (°C)') + xlab('') +
  geom_ribbon(data = uncertainty, aes(x=Value,ymin=ymin,ymax=ymax), fill='grey70' ) +
  geom_col(aes(fill=Value > 0)) +
  scale_fill_brewer(palette = 'Set1') +
  scale_x_continuous(expand = expand_scale()) +
  theme_cowplot(26) + theme(legend.position = 'none')
```

---

# Perceived trends

```{r, fig.align='center', fig.width=14,fig.height=7, echo=FALSE, warning=FALSE}
periods = global.climate %>% mutate(
  grp = cut(Year, c(1879,1957,1968,1969,1977,1980,1997,2001,2012,2015,2020)),
  grp2 = fct_other(grp, keep = levels(grp)[c(2,4,6,8,10)], other_level = 'other')
)
ggplot(global.climate, aes(Year,Value)) + 
  ylab('deviation from 1990 (°C)') + xlab('') +
  geom_line(size=1) +
  coord_cartesian(xlim = c(1950,2020)) +
  theme_cowplot(26) + theme(legend.position = 'none') +
  geom_smooth(size=2,
    data=filter(periods, grp2 != 'other'),
    aes(color=grp2), method = 'lm', se = FALSE) +
  scale_color_brewer(palette = 'Set1')
```

---

# Long-term trend

```{r, fig.align='center', message=FALSE, fig.width=14,fig.height=7, echo=FALSE, warning=FALSE}
ggplot(global.climate, aes(Year,Value)) + 
  ylab('deviation from 1990 (°C)') + xlab('') +
  geom_line(size=2, color='purple') +
  scale_x_continuous(expand = expand_scale()) +
  geom_smooth(color='yellow') +
  theme_cowplot(26) + theme(legend.position = 'none')
```


---

## Patterns

$$1, 4, 7, 10, \ldots$$

--

$$1, 3, 6, 10, 15, \ldots$$

--

$$0, 1, 1, 2, 3, 5, 8, \ldots$$

--

$$8, 5, 4, 9, 1, \ldots$$

???

* 13
* 21
* 13
* 7, 6, 10, 3




---

# Forecasting wrap-up

* How do we make good forecasts?

* What causes us to make bad predictions?

* How can you tell if predictions are good or bad?

---

# Important point from last week

* The need to learn more (how was the survey conducted, how many, who was in the sampling frame, are the analysts trustworthy)