---
title: "Week 7"
author: "DJM"
date: "15 February 2020"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: [default, rladies, rladies-fonts, extra.css]
    nature:
      beforeInit: ["macros.js", "https://platform.twitter.com/widgets.js"]
      titleSlideClass: [center,middle, inverse, bigger]
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
---


# Today's news


[Chocolate and Cognition](https://www.nejm.org/doi/full/10.1056/NEJMon1211064)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, autodep=TRUE,
                  cache=TRUE)
library(tidyverse)
library(ggforce)
library(cowplot)
```


---

# For Thursday

.center[.middle[
# Bring a laptop
]]

---

# Conditional probability

.pull-left[
* Monty Hall is an example of this (a deliberately challenging one)

* S is called the __sample space__, the collection of possible outcomes

* A and B are __events__, possible outcomes of interest

* Examples:

]

.pull-right[
```{r}
df = tibble(x = c(-1.5,1), y=c(0,0), lab = LETTERS[1:2], rad=c(2,1.5))
g=ggplot(df, aes(x0=x,y0=y,r=rad,fill=lab)) + geom_circle(alpha=.6) + 
  coord_fixed() + theme_void(24) + scale_fill_brewer(palette='Dark2') +
  theme(panel.border = element_rect(fill=NA,color="black",size=2),
        legend.position = 'none') +
  annotate('text', x=-3.5, y=1.75,label='S',size=16)
g +  geom_text(aes(x,y,label=lab),size=16)
```
]

---

# Unions, intersections, complements

$$A \cap B = A \textrm{ and } B$$


```{r, fig.align='center'}
g + annotate("text", x=0, y=0, label="A~intersect()~B", parse=TRUE,size=8)
```

---

# Unions, intersections, complements

$$A \cup B = A \textrm{ or } B$$


```{r, fig.align='center'}
ggplot(df, aes(x0=x,y0=y,r=rad)) + geom_circle(fill='yellow',color='yellow') + 
  coord_fixed() + theme_void(24) + scale_fill_brewer(palette='Dark2') +
  theme(panel.border = element_rect(fill=NA,color="black",size=2),
        legend.position = 'none') +
  annotate('text', x=-3.5, y=1.75,label='S',size=16) + 
  annotate("text", x=0, y=0, label="A~union()~B", parse=TRUE,size=16)
```

---

# Unions, intersections, complements

$$(A \cup B)^c = \textrm{neither } A \textrm{ nor } B$$


```{r, fig.align='center'}
ggplot(df, aes(x0=x,y0=y,r=rad)) + geom_circle(fill='white',color='white') + 
  coord_fixed() + theme_void(24) + scale_fill_brewer(palette='Dark2') +
  theme(panel.border = element_rect(fill=NA,color="black",size=2),
        legend.position = 'none', panel.background = element_rect(fill='yellow')) +
  annotate('text', x=-3.5, y=1.75,label='S',size=16) + 
  annotate("text", x=0, y=0, label="(A~union()~B)^c", parse=TRUE,size=16)
```

---

# Conditional on A

.big[
$$P(B\ |\ A) = \textrm{Given that B happens, what is the probabality that A happens?}$$
]

```{r, fig.align='center'}
ggplot(df, aes(x0=x,y0=y,r=rad,fill=lab)) + geom_circle(alpha=.6) + 
  coord_fixed(xlim=c(-3.6,.6)) + theme_void(24) +
  scale_fill_brewer(palette='Dark2') +
  theme(legend.position = 'none') +
  annotate("text", x=0, y=0, label="B | A", parse=FALSE,size=16) +
  annotate("text", x=-1.5,y=0,label="S %=>% A", parse=TRUE,size=16)
```

---

# Conditional on B

.big[
$$P(A\ |\ B) = \textrm{Given that A happens, what is the probabality that B happens?}$$
]

```{r, fig.align='center'}
ggplot(df, aes(x0=x,y0=y,r=rad,fill=lab)) + geom_circle(alpha=.6) + 
  coord_fixed(xlim=c(-.5,2.5)) + theme_void(24) +
  scale_fill_brewer(palette='Dark2') +
  theme(legend.position = 'none') +
  annotate("text", x=0, y=0, label="A | B", parse=FALSE,size=16) +
  annotate("text", x=1.5, y=0,label="S %=>% B", parse=TRUE,size=16)
```

---

# Notes

* Compare, $P(A \cap B)$ with $P(A\ |\ B)$ and $P(B\ |\ A)$

* Each depends on $P(A \cap B)$

* But it depends on "relative to what"

--

$$P(A\cap B) = \frac{P(A\cap B)}{P(S)}$$
$$P(A\ |\ B) = \frac{P(A\cap B)}{P(B)}$$
$$P(B\ |\ A) = \frac{P(A\cap B)}{P(A)}$$